---
title: "5. Uniformity Tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Uniformity Tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we show how to detect the presence of random responses using the uniformity test, which shows that given respondent-level item order randomization, recorded responses will follow a uniform distribution in the absence of random responses.

```{r setup}
library(rankingQ)
```

## Recorded Responses

In an earlier vignette, we have shown how to perform the bias correction on ranking questions using the anchor question. But how do we know if there is a need for bias correction in the first place?

Distribution of rankings of recorded responses can show whether there are indeed random responses. First, what are recorded responses?

## Uniformity Test

In this example, given there are 4! = 24 possible ways to rank the items, the proportion of recorded responses for each possible ranking should converge to 1/24 = 0.0417, in the absence of random responses. 

## Validity of the Anchor Question

Checking for uniformity also validates the usage of the anchor question. 


