---
title: "3. Analysis of Bias-corrected Ranking Data"
author: "Yuki Atsusaka and Seo-young Silvia Kim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The estimated weights from `imprr_weights` can be used to perform any analyses. For example, to estimate the average rank of party, one can leverage linear regression as follows:

```{r}
library(rankingQ)
library(estimatr)

data(identity_w)
lm_robust(party ~ 1, data = identity_w, weights = w) |> tidy()
```

While this illustrative example provides a valid point estimate, its confidence interval does not account for the estimation uncertainty around the estimated weights. Thus, in practice, `imprr_weights` must be used along with bootstrapping, such as the one available in `rsample` ([example](https://declaredesign.org/r/estimatr/articles/estimatr-in-the-tidyverse.html#bootstrap-using-rsample)).
