---
title: "1. Overview"
author: "Yuki Atsusaka and Seo-young Silvia Kim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the main `rankingQ` workflow using the
`identity` dataset. The package estimates various ranking-based
quantities from ranking data. It also allows researchers to correct for
measurement error caused by inattentive survey respondents.

```{r setup, message = FALSE, warning = FALSE}
library(rankingQ)
library(dplyr)

data(identity)
```

## Example Data

The `identity` dataset contains Americans' rankings about four sources
of identity and an anchor-ranking question with a known correct
ordering. The four items include political party, religion, gender, and
race. The key theoretical concept is *relative partisanship*—the extent
to which people prioritize partisanship over other sources of identity.

```{r}
identity |> 
  rename(party = app_identity_1,
         religion = app_identity_2,
         gender = app_identity_3,
         race = app_identity_4) |>
  select(
    app_identity, 
    party, religion, gender, race
  ) |>
  head()
```

It also includes the survey weight `s_weight`.

```{r}
identity |>
  select(
    s_weight,
    app_identity,
    starts_with("app_identity_"),
    anc_correct_identity
  ) |>
  head()
```

The `app_identity` columns describe the ranking question of interest,
while `anc_correct_identity` indicates whether each respondent answered
the anchor question correctly.

## Estimate Ranking-Based Quantities

We begin by estimating various ranking-based quantities with no bias
correction. We include survey weights via the `weight` argument.

The `imprr_direct` function **impr**ove **r**anking analysis directly by
estimating bias-corrected quantities of interest such as average ranks,
pairwise ranking probabilities, top-k probabilities, and marginal rank
probabilities.

```{r}
out_direct <- imprr_direct(
  data = identity,
  J = 4,
  main_q = "app_identity",
  weight = "s_weight"
)
```

The first output summarizes the estimated proportion of random
responses. As expected, no random response was detected (no bias
correction).

```{r}
out_direct$est_p_random
```

The second output contains several corrected ranking-based quantities of
interest.

```{r}
out_direct$results
```

## Direct Bias Correction

Now, we compute the above quantities by detecting random responses and
applying bias correction. The `imprr_direct` function takes another
argument `anc_correct`, which is a dummy variable that takes 1 if a
respondent has the right answer for the anchor question and 0 otherwise.

```{r}
out_direct <- imprr_direct(
  data = identity,
  J = 4,
  main_q = "app_identity",
  anc_correct = "anc_correct_identity",
  weight = "s_weight"
)
```

Now, the function returns the estimated proportion of random responses.
We find that about 35\\% [31%-40%] of respondents—a sizable share of
data—seem to provide random responses.

```{r}
out_direct$est_p_random
```

Finally, we obtain bias-corrected estimates of various quantities of
interest. Here, we focus on average rank. The estimated average ranks
are based on our plug-in bias-corrected estimator.

```{r}
out_direct$results |>
  filter(qoi == "average rank")
```

## Inverse-Probability Weighting

Instead of directly correcting for bias, the `imprr_weights` function
produces respondent-level (bias-correction) weights that can be used in
downstream analyses.

```{r}
out_weights <- imprr_weights(
  data = identity,
  J = 4,
  main_q = "app_identity",
  anc_correct = "anc_correct_identity",
  weight = "s_weight"
)
```

One output gives the bias-correction weight assigned to each possible
ranking profile.

```{r}
out_weights$rankings |>
  select(ranking, weights) |>
  head()
```

The respondent-level output keeps the original data and appends a
`weights` column along with a unified `ranking` column. To combine our
bias-correction weights with survey weights, users can simply create a
new variable that multiplies both weights.

```{r}
out_weights$results |>
  select(weights, s_weight, app_identity, ranking) |>
  mutate(joint_weight = weights * s_weight) |>
  head()
```

## Using the IPW Weights

The IPW-adjusted respondent-level data can be passed to downstream
helpers such as `avg_rank`.

```{r}
items_df <- data.frame(
  variable = paste0("app_identity_", 1:4),
  item = c("Party", "Religion", "Gender", "Race")
)

ipw_df <- out_weights$results |>
  mutate(joint_weight = weights * s_weight)

avg_rank(
  ipw_df,
  items = items_df,
  weight = "joint_weight",
  raw = FALSE
)
```

## Next Steps

The remaining vignettes go into more detail on specific parts of the
workflow:

1.  `2. Data` describes our example dataset.
2.  `3. Methods` covers the correction methods in more depth.
3.  `4. Analysis` shows downstream analysis with corrected weights.
4.  `5. Visualization` introduces the plotting helpers.
5.  `6. Test` covers diagnostics when anchor questions are unavailable
    or need validation.
