---
title: "1. Getting Started"
author: "Yuki Atsusaka and Seo-young Silvia Kim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the main `rankingQ` workflow using the
`identity` dataset. The package is designed for ranking questions that
may contain random responses and uses an anchor-ranking question to
estimate and correct the resulting measurement error.

```{r setup, message = FALSE, warning = FALSE}
library(rankingQ)
library(dplyr)

data(identity)
```

## The Example Data

The `identity` dataset contains a main ranking question about four
sources of identity and an anchor-ranking question with a known correct
ordering. It also includes the survey weight `s_weight`.

```{r}
identity |>
  select(
    s_weight,
    app_identity,
    starts_with("app_identity_"),
    anc_correct_identity
  ) |>
  head()
```

The `app_identity` columns describe the ranking question of interest,
while `anc_correct_identity` indicates whether each respondent answered
the anchor question correctly.

## Direct Bias Correction

The `imprr_direct` function directly estimates bias-corrected quantities
of interest such as average ranks, pairwise ranking probabilities,
top-k probabilities, and marginal rank probabilities.

```{r}
out_direct <- imprr_direct(
  data = identity,
  J = 4,
  main_q = "app_identity",
  anc_correct = "anc_correct_identity",
  weight = "s_weight"
)
```

The first output summarizes the estimated proportion of random
responses.

```{r}
out_direct$est_p_random
```

The second output contains several corrected ranking-based quantities of
interest. For a first pass, average ranks are often the easiest place to
start.

```{r}
out_direct$results |>
  filter(qoi == "average rank")
```

## Inverse-Probability Weighting

The `imprr_weights` function instead produces respondent-level weights
that can be used in downstream analyses.

```{r}
out_weights <- imprr_weights(
  data = identity,
  J = 4,
  main_q = "app_identity",
  anc_correct = "anc_correct_identity",
  weight = "s_weight"
)
```

One output gives the bias-correction weight assigned to each possible
ranking profile.

```{r}
out_weights$rankings |>
  select(ranking, weights) |>
  head()
```

The respondent-level output keeps the original data and appends a
`weights` column along with a unified `ranking` column.

```{r}
out_weights$results |>
  select(weights, s_weight, app_identity, ranking) |>
  head()
```

## Using the IPW Weights

The IPW-adjusted respondent-level data can be passed to downstream
helpers such as `avg_rank`.

```{r}
items_df <- data.frame(
  variable = paste0("app_identity_", 1:4),
  item = c("Party", "Religion", "Gender", "Race")
)

avg_rank(
  out_weights$results,
  items = items_df,
  weight = "weights",
  raw = FALSE
)
```

## Next Steps

The remaining vignettes go into more detail on specific parts of the
workflow:

1. `2. Basic Setup` describes the expected input data structure.
2. `3. Correcting Bias in Ranking Data` covers the correction methods in
   more depth.
3. `4. Analysis of Bias-corrected Ranking Data` shows downstream
   analysis with corrected weights.
4. `5. Visualizing Rankings` introduces the plotting helpers.
5. `6. Uniformity Tests` covers diagnostics when anchor questions are
   unavailable or need validation.
6. `7. Simulating Ranking Data` describes the data-generation helper
   `rpluce`.
