---
title: "1. Overview"
author: "Yuki Atsusaka and Seo-young Silvia Kim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette introduces the main `rankingQ` workflow using the
`identity` dataset. The package estimates various ranking-based
quantities from ranking data. It also allows researchers to correct for
measurement error caused by inattentive survey respondents.

```{r setup, message = FALSE, warning = FALSE}
library(rankingQ)
library(dplyr)
```

## Example Data

The `identity` dataset contains data on how Americans rank four sources
of identity that are central to American politics. The four items
include political party, religion, gender, and race. The key theoretical
concept is *relative partisanship*—the extent to which people prioritize
partisanship over other sources of identity.\
\
Below, the `app_identity` column stores the full ranking profile.
Similarly, the item columns (`party`, `religion`, `gender`, `race`)
store the marginal ranks.

```{r}
data(identity)

identity |>
  select(
    app_identity,
    party, religion, gender, race
  ) |>
  head()
```

It also includes the survey weight `s_weight`. Additionally,
`anc_correct_identity` is the binary variable for whether respondents
provide the correct answer to the anchor ranking question.

```{r}
identity |>
  select(
    app_identity,
    party, religion, gender, race,
    s_weight,
    anc_correct_identity
  ) |>
  head()
```

Substantively, [Atsusaka and Kim
(2025)](https://doi.org/10.1017/pan.2024.33) are interested in the
extent to which *political party* is important when it comes to people's
multidimensional identity.

## Estimate Ranking-Based Quantities

Now, let us demonstrate how to compute various ranking-based quantities
based on the data. We begin by estimating such quantities with no bias
correction. To make it realistic, however, we include survey weights via
the `weight` argument.

### Input

The `imprr_direct` function **impr**ove **r**anking analysis
**direct**ly (as in a plug-in way) by estimating bias-corrected
quantities of interest. These quantities include average ranks, pairwise
ranking probabilities, top-k probabilities, and marginal rank
probabilities. 

Here, `main_q` argument takes a vector of all items in
the choice set.

```{r}
out_direct <- imprr_direct(
  data = identity,
  J = 4,
  main_q = c("party", "religion", "gender", "race"),
  weight = "s_weight"
)
```

### Output

`imprr_direct` returns two lists as an output.\
\
The first output summarizes the estimated proportion of random
responses. As expected, no random response was detected (no bias
correction).

```{r}
out_direct$est_p_random
```

The second output contains several corrected ranking-based quantities of
interest.

```{r}
out_direct$results
```

Researchers can examine any classes of ranking-based quantities. For example, they can filer out the results by focusing on top-k ranking probabilities:

```{r}
out_direct$results |>
  filter(qoi == "top-k ranking")
```



## Apply Bias Correction

Now, suppose that we are concerned that the original data contain random responses or satisficing answers. We worry that such responses would introduce measurement error to our data.

The `rankingQ` package offers two ways to address such concern. Both approaches allow us to compute bias-corrected estimates of our quantities of interest.


### Plug-in Bias-corrected Estimator

The first approach is to account for the proportion of random responses and directly bias correct our estimates.

To estimate the proportion of random responses, [Atsusaka and Kim
(2025)](https://doi.org/10.1017/pan.2024.33) advocated using an anchor ranking question: an auxiliary ranking question whose correct answer is known to researchers and respondents. To precisely measure the level of satisficing responses in the target ranking question, we recommend that researchers ask the anchor question right before or after the primary ranking question.

The `imprr_direct` function takes an additional
argument `anc_correct`, which is a dummy variable that takes 1 if a
respondent has the right answer for the anchor question and 0 otherwise.

```{r}
out_direct <- imprr_direct(
  data = identity,
  J = 4,
  main_q = c("party", "religion", "gender", "race"),
  weight = "s_weight",
  anc_correct = "anc_correct_identity" # additional input
)
```

In this example, the function returns the estimated proportion of random responses. We find that about 35\\% [31%-40%] of respondents—a sizable share of
data—seem to provide random responses.

```{r}
out_direct$est_p_random
```

Our bias-corrected estimates are available in `results`. Here, we focus on average rank. Again, the estimated average ranks
are based on our plug-in bias-corrected estimator.

```{r}
out_direct$results |>
  filter(qoi == "average rank")
```

#### Using Attention Checks or Other Methods

In some applications, random responses may also be detected by other methods, including attention checks, screener questions, factual manipulation checks, and response time.

**Our package can accommodate these alternative methods with no problems.**

The only change is to use an alternative argument `p_random` and specify the estimated proportion of random responses directly.

```{r}
out_alternative <- imprr_direct(
  data = identity,
  J = 4,
  main_q = c("party", "religion", "gender", "race"),
  weight = "s_weight",
  p_random = 0.5 # estimated proportion of random responses
)
```

By definition, `est_p_random` returns the input value:

```{r}
out_alternative$est_p_random
```

The output format stays the same as before.

```{r}
out_alternative$results |>
  filter(qoi == "average rank")
```


### Inverse-Probability Weighting

The second approach is to estimate bias-correction weights and use the weights in downstream analyses.

The key idea is that some rankings are oversampled and others are undersampled due to measurement error. Thus, for rankings that are artificially overrepresented, we want to down weight them. For rankings that are underrepresented, we want to weight them up.

This is known as the inverse-probability weighting.

For this strategy, the `imprr_weights` function allows us to estimate bias-correction **weights** for each survey respondent.

```{r}
out_weights <- imprr_weights(
  data = identity,
  J = 4,
  main_q = c("party", "religion", "gender", "race"),
  weight = "s_weight",
  anc_correct = "anc_correct_identity", # additional input
)
```

What this approach does is to assign a bias-corrected weight to each possible ranking profile. The `rankings` list contains the comprehensive list of rankings with bias-correction weights. 

```{r}
out_weights$rankings |>
  select(ranking, weights) |>
  head()
```

The respondent-level output keeps the original data and appends a
`weights` column along with a unified `ranking` column.

In many cases, we wish to account for two types of weights: survey weights and bias-correction weights. We can do so by multiplying both weights to create a single (joint) weight variable. To do so, users can simply create a
new variable that multiplies both weights.

```{r}
out_weights$results |>
  select(weights, s_weight, app_identity, ranking) |>
  mutate(joint_weight = weights * s_weight) |>
  head()
```


The IPW-adjusted respondent-level data can be passed to any downstreatm analyses, including descriptive and regression analyses. 

We demonstrate it by using our in-house helper function `avg_rank`.

```{r}
items_df <- data.frame(
  variable = c("party", "religion", "gender", "race"),
  item = c("Party", "Religion", "Gender", "Race")
)

ipw_df <- out_weights$results |>
  mutate(joint_weight = weights * s_weight)

avg_rank(
  ipw_df,
  items = items_df,
  weight = "joint_weight",
  raw = FALSE
)
```

## Next Steps

The remaining vignettes go into more detail on specific parts of the
workflow:

1.  `2. Data` describes our example dataset.
2.  `3. Methods` covers the correction methods in more depth.
3.  `4. Analysis` shows downstream analysis with corrected weights.
4.  `5. Visualization` introduces the plotting helpers.
5.  `6. Test` covers diagnostics when anchor questions are unavailable
    or need validation.
