---
title: "3. Analysis of Bias-corrected Ranking Data"
author: "Yuki Atsusaka and Seo-young Silvia Kim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Analysis of Bias-corrected Ranking Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message=FALSE, warning=FALSE}
library(rankingQ)
library(estimatr)
data(identity_w)
```

The estimated weights from `imprr_weights` can be used to perform any analyses. For example, to estimate the average rank of party, one can leverage linear regression as follows:

```{r}
lm_robust(
  app_identity_1 ~ 1,
  data = identity_w,
  weights = identity_w$weights
) |>
  tidy()
```

While this illustrative example provides a valid point estimate, its confidence interval does not account for the estimation uncertainty around the estimated weights. Thus, in practice, `imprr_weights` must be used along with bootstrapping, such as the one available in `rsample` ([example](https://declaredesign.org/r/estimatr/articles/estimatr-in-the-tidyverse.html#bootstrap-using-rsample)).

## Computing Average Ranks

The `avg_rank` function provides a convenient way to compute average ranks for all items:

```{r}
# Raw average ranks (without bias correction)
avg_rank(identity_w,
  rankings = "app_identity",
  items = c("Party", "Religion", "Gender", "Race")
)
```

For bias-corrected estimates using IPW weights, use the marginal ranking columns:

```{r}
# IPW-corrected average ranks
items_df <- data.frame(
  variable = paste0("app_identity_", 1:4),
  item = c("Party", "Religion", "Gender", "Race")
)
avg_rank(identity_w, items = items_df, weight = "weights", raw = FALSE)
```
